---
title: "Ari C"
format: html
---

```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)

library(shiny)

library(tibble)

library(viridisLite)

```

```{r}
#| echo: false
#| message: false
#| warning: false
#| results: hide

read_csv("../../data/wcmatches.csv")
```


```{r}
unemployment_data <- read_csv("../../data/Brazil-Unemployment-Rate-Unemployment-Rate-2025-04-16-11-18.csv") |>
  slice_head(n = 30) |>
  slice_tail(n = 11) |>
  rename(Year = ...1)
 
```

```{r}
ggplot(unemployment_data, aes(x = Year, y = `Unemployment Rate`)) +
  geom_jitter() +
  geom_line() +
  scale_x_continuous(breaks = c(2010:2020)) 
```
# Key Takeaways

- The unemployment rate steadily decreased from 2010-2014 going from 8.42% to 6.76% in 2014, the year of the world cup
- After the world cup, from 2014-2020, the unemployment skyrocketed from 6.76% to 13.70%, an more than double increase in the span of 6 years

# Key Questions

- Why was the unemployment rate decreasing from 2010-2014, and what caused the unemployment rate to increase after 2016?

- Is the world cup a key attribute to the increase in the unemployment rate, and/or are there other factors that contributed to this increase in umployment?



Inflation Rate:

```{r}
inflation_data <- read_csv("~/Documents/Brazil Inflation data.csv")
```

```{r}
inflation_data <- inflation_data|>
  rename(year = observation_date, inflation_rate = FPCPITOTLZGBRA)
```


```{r}
ggplot(inflation_data, aes(x = year, y= inflation_rate, fill = inflation_rate)) +
 geom_bar(stat = "identity") +
   scale_fill_gradient(low = "lightgrey", high = "lightgrey") +
  labs(
    title = "Brazil's Inflation Rate",
    caption = "Retrieved from FRED, Federal Reserve Bank of St. Louis",
    y = "Inflation Rate",
    x = "Year",
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0, size = 5))
```
```{r}
 "Unemployment rate in Brazil (2010-2020)"

ggplot(unemployment_data, aes(x = Year, y= `Unemployment Rate`, fill = `Unemployment Rate`)) +
  geom_bar(stat = "identity") +
   scale_fill_gradient(low = "lightgrey", high = "lightgrey") +
  scale_x_continuous(breaks = c(2010:2020)) +
  labs(
    title = "Unemployment rate in Brazil (2010-2020)" ,
    y = "Unemployment Rate (%)",
    caption = "Retrieved from FRED, Federal Reserve Bank of St. Louis"
    ) +
  theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(hjust = 0, size = 5))
```

