---
title: "PJ P"
format: html
---

```{r}
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(scales)
library(readr)
```


```{r}
gdp_data <- read_csv("../../data/brazil_econ 2/API_NY.GDP.MKTP.CD_DS2_en_csv_v2_26433.csv", skip = 4)

```


```{r}
brazil_gdp <- gdp_data %>%
  filter(`Country Name` == "Brazil") %>%
  select(`Country Name`, starts_with("19"), starts_with("20")) %>%
  pivot_longer(
    cols = -`Country Name`,
    names_to = "Year",
    values_to = "GDP"
  ) %>%
  mutate(Year = as.numeric(Year))

```

```{r}
ggplot(brazil_gdp, aes(x = Year, y = GDP)) +
  geom_line(color = "blue", size = 1.2) +
  scale_y_continuous(labels = scales::dollar_format(scale = 1e-9, suffix = "B")) +
  labs(
    title = "Brazil's GDP Over Time",
    x = "Year",
    y = "GDP (USD, Billions)"
  ) +
  theme_minimal()

```

```{r}

# Extract 2014 GDP value
gdp_2014 <- brazil_gdp %>% filter(Year == 2014)

ggplot(brazil_gdp, aes(x = Year, y = GDP)) +
  geom_line(color = "blue", size = 1.2) +
  geom_point(data = gdp_2014, aes(x = Year, y = GDP), color = "red", size = 3) +
  geom_text(
    data = gdp_2014,
    aes(label = paste0("2014: $", round(GDP / 1e9, 1), "B")),
    vjust = -1, hjust = 0.5,
    color = "red"
  ) +
  scale_y_continuous(labels = dollar_format(scale = 1e-9, suffix = "B")) +
  labs(
    title = "Brazil's GDP Over Time (Highlighted: 2014)",
    x = "Year",
    y = "GDP (USD, Billions)"
  ) +
  theme_minimal()

```
The visualization shows how Brazil's GDP was higher before the world cup most likely because they knew they would be hosting it and the government was spending money to get ready for it and then during the world cup it was high because they were bringing in a lot of money hosting it. However after the world cup the GDP dropped by a good amount most likely because now the government wasn't spendning money or bringing in money from the world cup so it dropped.

```{r}
summary_stats <- brazil_gdp %>%
  summarize(
    Years_Available = sum(!is.na(GDP)),
    First_Year = min(Year[!is.na(GDP)]),
    Last_Year = max(Year[!is.na(GDP)]),
    Min_GDP = min(GDP, na.rm = TRUE),
    Max_GDP = max(GDP, na.rm = TRUE),
    Mean_GDP = mean(GDP, na.rm = TRUE),
    Median_GDP = median(GDP, na.rm = TRUE),
    SD_GDP = sd(GDP, na.rm = TRUE)
  )

# View the summary
print(summary_stats)

```

```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(summarytools)


gdp_long <- gdp_data %>%
  select(`Country Name`, starts_with("19"), starts_with("20")) %>%
  pivot_longer(
    cols = -`Country Name`,
    names_to = "Year",
    values_to = "GDP"
  ) %>%
  mutate(Year = as.numeric(Year))

```

```{r}

# Summary of GDP across all countries and years
dfSummary(gdp_long$GDP)

```

```{r}
gdp_by_country <- gdp_long %>%
  group_by(`Country Name`) %>%
  summarize(
    start_year = min(Year, na.rm = TRUE),
    end_year = max(Year, na.rm = TRUE),
    num_years = sum(!is.na(GDP)),
    min_gdp = min(GDP, na.rm = TRUE),
    max_gdp = max(GDP, na.rm = TRUE),
    mean_gdp = mean(GDP, na.rm = TRUE),
    median_gdp = median(GDP, na.rm = TRUE),
    sd_gdp = sd(GDP, na.rm = TRUE),
    total_gdp = sum(GDP, na.rm = TRUE)
  ) %>%
  arrange(desc(mean_gdp))  # Optional: sort by average GDP
```

```{r}
head(gdp_by_country, 10)

```

```{r}
View(gdp_by_country)

```





```{r}

library(tidyverse)

# Skip the first three rows to get to the real header row
tourism_data <- read_csv("../../data/API_ST.INT.RCPT.CD_DS2_en_csv_v2_19613.csv", skip = 3)

# Filter for Brazil and select only year columns
brazil_data <- tourism_data %>%
  filter(`Country Name` == "Brazil") %>%
  select(`Country Name`, matches("^\\d{4}$"))  # Select only year columns

# Reshape for plotting
brazil_long <- brazil_data %>%
  pivot_longer(
    cols = -`Country Name`,
    names_to = "Year",
    values_to = "Receipts"
  ) %>%
  mutate(
    Year = as.integer(Year),
    Receipts = as.numeric(Receipts) / 1e9  # Convert to billions
  ) %>%
  filter(Year >= 2010 & Year <= 2018) %>%
  mutate(
    Highlight = if_else(Year == 2014, "2014", "Other")  # Highlight 2014
  )

# Plot with line connecting all years and 2014 highlighted
ggplot(brazil_long, aes(x = Year, y = Receipts)) +
  geom_line(color = "black", size = 1.2) +  # Line connecting all years
  geom_point(aes(color = Highlight), size = 3) +  # Points with different color for 2014
  scale_color_manual(values = c("2014" = "#FCD116", "Other" = "#009B4D")) +  # Color 2014 differently
  labs(
    title = "Brazil: International Tourism Receipts (2010-2018)",
    x = "Year",
    y = "Receipts (billion US$)",
    color = "Year"
  ) +
  theme_minimal()


```

```{r}
ggplot(brazil_long, aes(x = Year, y = Receipts)) +
  geom_line(color = "black", size = 1.2) +  # Line connecting all years
  geom_point(aes(color = Highlight), size = 3) +  # Points with different color for 2014
  geom_point(data = subset(brazil_long, Year == 2016), 
             aes(x = Year, y = Receipts), 
             color = "red", size = 4) +  # Red dot for 2016
  geom_text(data = subset(brazil_long, Year == 2016), 
            aes(label = "Olympics"), 
            vjust = -1, color = "red", size = 4.5) +  # Label for 2016
  scale_color_manual(values = c("2014" = "#FCD116", "Other" = "#009B4D")) +  # Color 2014 differently
  labs(
    title = "Brazil: International Tourism Receipts (2010-2018)",
    x = "Year",
    y = "Receipts (billion US$)",
    color = "Year"
  ) +
  theme_minimal()

```

```{r}
brazil_long$Highlight <- ifelse(brazil_long$Year == 2014, "World Cup",
                          ifelse(brazil_long$Year == 2016, "Olympics", "Other"))

ggplot(brazil_long, aes(x = Year, y = Receipts)) +
  geom_line(color = "black", size = 1.2) +  # Line connecting all years
  geom_point(aes(color = Highlight), size = 3) +  # Points with different color for 2014
  geom_point(data = subset(brazil_long, Year == 2016), 
             aes(x = Year, y = Receipts), 
             color = "red", size = 4) +  # Red dot for 2016
  geom_text(data = subset(brazil_long, Year == 2016), 
            aes(label = "Olympics"), 
            vjust = -1, color = "red", size = 4.5) +  # Label for 2016
  geom_text(data = subset(brazil_long, Year == 2014), 
            aes(label = "World Cup"), 
            vjust = -1, color = "#FCD116", size = 4.5) +  # Label for 2014
  scale_color_manual(values = c("2014" = "#FCD116", "Other" = "#009B4D")) +  # Color 2014 differently
  labs(
    title = "Brazil: International Tourism Receipts (2010-2018)",
    x = "Year",
    y = "Receipts (billion US$)",
    color = "Year"
  ) +
  theme_minimal()

```



